---
title: "Practical Machine Learning - Exercise type prediction project"
author: "Martin Bielik"
date: "October 22, 2017"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data cleaning - pre-processing

At first, I reduce the complexity of prediction problem by elliminating variables with missing or NA values.
The Participant names and exercise time stamps are removed as well

```{r read, echo=FALSE, cache=TRUE}
#training <- read.csv("I:/OneDrive/03_Data/01_Courses/Data Science Specialization/Practical Machine Learning/pml-training.csv")
#testing <- read.csv("I:/OneDrive/03_Data/01_Courses/Data Science Specialization/Practical Machine Learning/pml-testing.csv")
training <- read.csv("C:/Users/marti/OneDrive/03_Data/01_Courses/Data Science Specialization/Practical Machine Learning/pml-training.csv")
testing <- read.csv("C:/Users/marti/OneDrive/03_Data/01_Courses/Data Science Specialization/Practical Machine Learning/pml-testing.csv")

set.seed(988)
```

```{r pre-processing}
variable.CullBlank <- as.logical(training[1,] == "")
variable.CullNA <- is.na(variable.CullBlank)
variable.cull <- !variable.CullNA & !variable.CullBlank

variable.cull[1:5] <- F

trainingProc <- training[, variable.cull]
testingProc <- testing[, variable.cull]
testingProc <- testingProc[,1:(ncol(testingProc)-1)]
```

The final training data set was redced from the original `r ncol(testing)` to `r ncol(testingProc)`.

## Training

SInce the goal of this porject is prediction and not the ability to interpret and understand the link between sensor data and exercise quality, I adopted one of the most effective and universal machine learning algorithms - Random forest.

The out of the sample error was estimated by 5 fold cross-validation.

###Side note: 
The Random Fordest training was implemented as parallel computation.

```{r training, echo=TRUE, cache=TRUE, warning=FALSE}
library(caret)
library(doParallel)
library(parallel)

cluster <- makeCluster(detectCores() - 1)
registerDoParallel(cluster)

fitControl <- trainControl(method = "cv",
                           number = 5,
                           allowParallel = TRUE)
modFit.RF <- train(classe ~ ., data = trainingProc, method = "rf", trControl = fitControl)

stopCluster(cluster)
registerDoSEQ()
```
## Final model
```{r echo=FALSE}
modFit.RF$finalModel
CM <- confusionMatrix(modFit.RF)
```

The estimated out of sample model accuracy is `r sum(diag(CM$table))`%

## Predicting on test set

```{r predicting, echo=FALSE}
pred <- predict(modFit.RF, testingProc)
data.frame("ID-Question" = 1:20, "Prediction" = pred)
```

After inserting the prediction into the final quiz, all 20 predictions were marked as corret resulting into the accuracy of 100%
